### 2019-09-10 paper/unstructured.Rmd
fst intro

---
output: 
  github_document:
    toc: true
author: Alec Robitaille, Quinn M.R. Webber, Julie Turner and Eric Vander Wal
date: "-- `r Sys.Date()`"
---

#### From issue organisers

* Appreciate ambition to link LE and SNA.
* Like the idea of reviewing different kinds of spatial data relevant to ML framework
* Hope simulations will be as relevant to movement and SNA data typically collected
* Manuscript will be relevant to both theoreticians and empiricists

## Keywords

1. spatial scale
1. social systems
1. multilayer networks
1. landscape ecology
1. network theory

## Boxes / paper layout

* Abstract
* Intro
* Box 1: Mathematical + language definitions of multilayer networks
* Box 2: Types of spatiotemporally explicit data and considerations relevant to these e.g. fix rate/grouping with GPS and gambit of the group, etc (table)
* Spatial variables - example of GPS data, sampling landcover and networks across landcover classes?
* Temporal variables - example of GoG data, chunked seasonally maybe?
* Relevant multilayer metrics with above two? 
* Conclusion/fin