### 2019-09-11 paper/unstructured.Rmd
rm temporal

---
output: 
  github_document:
    toc: true
author: Alec Robitaille, Quinn M.R. Webber, Julie Turner and Eric Vander Wal
date: "-- `r Sys.Date()`"
---

#### From issue organisers

* Appreciate ambition to link LE and SNA.
* Like the idea of reviewing different kinds of spatial data relevant to ML framework
* Hope simulations will be as relevant to movement and SNA data typically collected
* Manuscript will be relevant to both theoreticians and empiricists

**Conservation implications**

## Keywords

1. spatial scale
1. social systems
1. multilayer networks
1. landscape ecology
1. network theory

## Skeleton
* Abstract
* Brief introduction
  * Networks, social networks, animal social systems 
  * Landscape ecology, scale
* Multilayer Networks
  * What is a ML network
  * What kinds of ML nets exist
  * Which are relevant to spatial variables
* Scale 
  * What is scale
  * In LE lit, how has scale been investigated
  * Elements of scale relevant to animal social systems, animal movement data, landcover data
* Box 1: Types of spatially explicit data and considerations relevant to these e.g. fix rate/grouping with GPS and gambit of the group, etc (table)
* Box 2: Mathematical + language definitions of multilayer networks, metrics and measures of ML networks that are relevant to spatial data
* Simulations
* Discussion/results
* Brief conclusion

---

From a sample of CZ papers:

* Abstract
* Materials and Methods
* Results
* Discussion

and sometimes
* Conservation implications
* Conclusion