### 2019-09-27 paper/table1.Rmd
rm old pander

---
output: 
  # word_document:
  #   reference_docx: template.docx
  github_document
author: Alec Robitaille, Quinn M.R. Webber, Julie Turner and Eric Vander Wal
date: "`r Sys.Date()`"
bibliography: ../references/refs.bib

---

### Table 1
In Table 1, we describe three types of multilayer networks that incorporate spatial variables. Multilayer networks are composed of layers, nodes and edges. Layers represent different types of relationships

The edges are often abstracted because they no longer in the sense that they represent a specific spatial and temporal context, which

[@dale2017applying]

Layers represent the different types of relationships, individuals or contexts of the social system. Nodes represent individual entities in the network, connected by some association or relationship with edges. Interlayer edges connect layers to other layers or individuals across layers. Intralayer edges connect individuals to other individuals within a network layer.

Relationships between individuals can be differentiated by their physical distance: communication or other high distance, grooming (physical contact), association (threshold distance)

Layers represent a type of edge (e.g.: grooming and association), entity (e.g.: male and female, or different species), or spatial context (e.g.: landcover type, patch).

Depending on the type of network, interlayer edges connect individuals to themselves across different contexts or to different individuals

multilayer networks with layers defined by spatially explicit locations, interspecific interactions, and community or subpopulation structure. 
Other types of networks, such as intercontexual networks or temporal networks, are likely only to incorporate spatial variables in a hybrid multilayer network and considerations for including those variables will be explored here

An animal's social network is multifaceted and complex, though this system is often examined

How are layers defined, nodes defined, inter layer edges

We list variables affected by spatial and social scale

We use a definition of spatial scale from LE,

social scale is

```{r, echo = FALSE, warning = FALSE}
library(data.table)

tab1 <- fread('table1.csv')

options(knitr.table.format = "markdown")

knitr::kable(tab1)
```