### 2019-09-30 paper/table1.Rmd
add wg notes

---
output: 
  word_document:
    reference_docx: template.docx
  # github_document
author: Alec Robitaille, Quinn M.R. Webber, Julie Turner and Eric Vander Wal
date: "`r Sys.Date()`"
bibliography: ../references/refs.bib
csl: vancouver.csl
---

<!--
*network of networks is just an intraspecific network* *interspecific networks are networks of networks*
-->

### Table 1
Multilayer networks are composed of layers, nodes and edges. Layers represent different types of relationships (e.g.: grooming or association), types of entities (e.g.: male and female, or different species), and spatial locations (e.g.: landcover type, patch) [@pilosof2017]. Nodes represent individual entities in the network, connected by some association or relationship with edges. Interlayer edges connect layers to other layers or individuals across layers. Intralayer edges connect individuals to other individuals within a layer. In Table 1, we describe three types of multilayer networks that incorporate spatial variables: spatially explicit networks, networks of networks and interspecific networks. <!-- not clear --> Other types of networks, such as intercontexual networks or temporal networks, are likely only to incorporate spatial variables in a hybrid multilayer network and relevant considerations will be detailed here.

<!-- 
highlight examples at the beginning? in the text?

Relationships between individuals can be differentiated by their physical distance: communication or other high distance, grooming (physical contact), association (threshold distance)

An animal's social network is multifaceted and complex, though this system is often examined

The edges are often abstracted because they no longer in the sense that they represent a specific spatial and temporal context, which

mv type of network across the top

what about intralayer social grain -->

```{r, echo = FALSE, warning = FALSE}
library(data.table)

tab1 <- fread('table1.csv')

options(knitr.table.format = "markdown")

knitr::kable(tab1)
```

# References