### 2019-12-08 paper/introduction.Rmd
first paragraph

---
output: github_document
bibliography: ../references/refs.bib
---

# Introduction

<!-- Landscape spatial scale -->
Our understanding of ecological patterns and processes are directly linked to the scale at which they are observed [@levin1992].

heteroneity of landscapes
within seemingly homogeneous patches, exist lower level processes including species interactions, resource availability, disease transfer [@legendre1993]

Animals may be influenced by processes at multiple scales

<!-- observational scales of landscapes do not necessarily correspond to the "scale of effect" of species-landscape relationships, or the scale at which there is the strongest relationship [@jackson2015]-->

the scales that influence animals and the patterns of their landscapes may be disjoint [@chave2013]. The effects of scale have been explicitly investigated across many ecological studies including habitat selection [@mayor2007], evaluating satellite imagery [@benson1995] and    of ecology including landscape ecology, movement ecology, earth observation [@benson1995], habitat selection[@mayor2007]

<!-- observers bias their perspective when selecting a scale,often unknowingly or without considering the implications [@allen2015].  -->

Scale can be considered for all measured quantities [@schneider2009] including spatial and temporal variables as well as domain specific <!--lead into social-->

<!-- Social proximity scale 
sociality exists 
-->
Social scale is the type of social relationship measured and how this relationship is defined [@farine2015]. If the greater range of a social relationship is the entire social system for an organism, a particular social scale may focus on only a limited type of relationship like grooming or aggression.

<!-- Regular networks
* SNA significant outcomes
-->

<!-- Multilayer networks
used across disciplines
-->
Multilayer networks explicitly consider social systems across temporal, spatial and social contexts [@pilosof2017a]. They are made up of multiple layers, each representing relationships, individuals, spatial areas or temporal windows [@kivela2014; @porter2018a]. <!--GENERALLY INTERESTING Recently used in animal behaviour studies, multilayer networks have the potential to better characterize complex social systems [@silk2018a; @finn2019].-->

For multilayer social networks, three types of scale are relevant: spatial, temporal and social scale.

<!-- Case study 
deer
fogo caribou
-->

integrating networks across scales
explore sensitivity of network measures to scale

---

Spatial scale is a central consideration in the way we understand landscape ecology and animal space use.

Despite the interplay between social networks and how animals use space, there remains an important biological and methodological gap in our understanding of how an animals perception of spatial scale, e.g., grain and extent,

We will simulate animal movement data coupled with emergent social networks on increasingly heterogenous landscapes across spatial scales. Furthermore, we review specific examples of spatially explicit data types to include in multilayer networks and explore the relevant nuances of these data. Our simulations will highlight the need to carefully select scales for measuring spatial processes and appropriately weight layers and edges representing different types of data. Effective integration of spatial processes, including biologically meaningful scales, into animal social networks is an emerging area of research. Our contribution will incorporate landscape ecology theory as a way to link different scales of social and spatial processes in a multilayer network. Based on social network and landscape ecology theory as well as our simulations we demonstrate potential for interconnectedness among individual animals and the landscapes they occupy.

---

potential for interconnectedness among individual animals and the landscapes they occupy.

Animal social systems are complex and multifaceted

traditionally, animal social systems are measured using direct observation, telemetry or pit tags
interactions, e.g. grooming, aggression, etc,  between individuals are quantified by simplified measures 
and often lacking any indication of the spatial context of the interaction

recently the multilayer has been used to

along with spatial variables, comes a large history of landscape ecology exploring and evaluating sensitivity to scale

here we'll use simulations to evaluate the impact of scale on social and spatial variables 
to better understand it's impact, implications in ML networks

# Multilayer Networks
Multilayer networks combine network layers with interlayer edges

Two main types of multilayer networks exist: multiplex networks and interconnected networks. Layers in multiplex networks consist of the same entities with each layer representing either a different type of interaction (as in multirelational networks) or a different time point or window (as in temporal networks). Layers in interconnected networks consist of the same or different entities with layers representing different subsystems (as in networks of networks), different types of entities (as in intercontextual networks), or different spatial locations (as in spatial networks) [@finn2019].

Of these, we have identified multirelational networks, networks of networks, intercontextual networks and spatial networks as likely to have spatial and scale considerations

Layers in multilayer networks can be
Animal social systems are

What is a network

How does a ML network build on that

Which can include spatial variables

Multifaceted nature of animal social systems 
many variables having a spatial component
including landcover and habitat type
fix rate and

spatial variables defined by

# Scale
perception of scale

What is scale

in LE lit, how has scale been investigated

Elements of scale relevant to animal social systems, animal movement data, landcover data

Grain is this, extent is this

How might scale influence results

What kinds of spatial scale are there

# References