### 2019-12-09 paper/introduction.Rmd
tidy + fin ml (for now)

---
output: github_document
bibliography: ../references/refs.bib
csl: cite.csl
---

# Introduction

<!-- ### Landscape spatial scale -->
Our understandings of ecological pattern and process are directly linked to the scale at which they are observed [@levin1992problem; @allen2015]. Animals and landscapes are influenced across scales by higher level processes, such as climate and productivity [@field2009], and lower level processes, such as species interactions and resource availability [@legendre1993; @chave2013]. The effect of scale has been explicity considered in ecology including in species distribution [@elith2009], habitat selection [@mayor2007] and food webs [@sugihara1989b]. Scale can be considered for all measured quantities [@schneider2009] including spatial and temporal variables as well as domain specific variables such as sociality.

<!-- observational scales of landscapes do not necessarily correspond to the "scale of effect" of species-landscape relationships, or the scale at which there is the strongest relationship [@jackson2015]-->

<!-- the scales that influence animals and the patterns of their landscapes may be disjoint [@chave2013].  -->

<!-- observers bias their perspective when selecting a scale,often unknowingly or without considering the implications [@allen2015].  -->

<!-- Spatial scale is a central consideration in the way we understand landscape ecology and animal space use.  -->

<!-- ### Social proximity scale  -->
Sociality, defined simply as the degree of shared space use among conspecifics [@tinbergen1951], exists across scales. For example, grooming and mating relationships requires close proximity between conspecifics and occurs over brief time periods whereas home range overlap occurs over larger spatial and temporal scales. Shared space is a requirement for social interaction, however, it remains unclear whether spatial proximity scales predictably with social interactions [@castles2014a; @farine2015a]. Social scale can be defined as the type of relationship observed [@farine2015a] and these relationships can be explicitly defined and measured in social network analyses [@farine2015b].

<!-- ### Regular networks
* SNA significant outcomes
-->

Rooted in mathematical graph theory, animal social networks measure the relationships of individuals and organization of social systems [@krause2009; @wey2008a; @croft2008]. Social network analyses provide insight into mating behaviours, interspecific interactions, and community structures [@pinter-wollman2014]. 
The structure of social networks influences population dynamics and evolutionary processes [@kurvers2014]. Nevertheless, social network analyses usually only consider a single measure of sociality, drastically simplifying the complexity of animal social systems [@finn2019].

<!-- Network theory and social network analysis allows one to scale across dyads, groups, etc from pairwise interactions to social units, networks etc[@krause2009] -->

<!-- infer process from large scale pattern [@krause2009] -->

<!-- ### Multilayer networks
used across disciplines
-->
Multilayer networks explicitly consider social systems across contexts [@pilosof2017a]. They are made up of multiple layers, each representing types of relationships or individuals, spatial areas and temporal windows [@kivela2014; @porter2018a]. Multilayer networks have been used to described multidimensional human social systems, complex transportation networks and the complete neural network of an organism [@boccaletti2014]. Recently, multilayer networks have been highlighted as a way to better characterize complex social systems in animal behaviour studies [@silk2018a; @finn2019].

<!-- For multilayer social networks, three types of scale are relevant: spatial, temporal and social scale. -->

<!-- ### Case study 
deer
fogo caribou
-->
We will use two case studies to explore the sensitivity of multilayer network measures across scales. One, using a population of caribou (*Rangifer tarandus*) on Fogo Island, Newfoundland, Canada with seasonal differences in resource availability and social association [@peignier2019]. The second using a small group of white-tailed deer (*Odocoileus virginianus*) in an urban environment **WHERE XYZ**. We will manipulate the spatial scale of landscapes, temporal windows of analysis and social scale of relationships, highlighting the importance of considering scale in multilayer networks.

# References