### 2019-12-11 scripts/explore-networkx.Rmd
from_pandas_dataframe -> from_pandas_edgelist

---
title: "explore networkx"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source('../scripts/1-single-scale.R')
```

Expectation is an edge list

```{python}
import pandas as pd
import networkx as nx

edges = pd.read_csv('../data/edgels.csv')
edges
G=nx.from_pandas_edgelist(edges, 'V1', 'V2', edge_attr='w')
```