### 2019-12-17 scripts/explore-networkx.Rmd
separate graphs by lc

---
title: "explore networkx"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source('../scripts/1-single-scale.R')
```

Expectation is an edge list

```{python}
import pandas as pd
import networkx as nx
import multinetx as ml

edges = pd.read_csv('../data/edgels.csv')

print(edges['relc'].unique())

G1 = nx.from_pandas_edgelist(edges[edges['relc'] == 1], 'V1', 'V2', edge_attr='w')
G2 = nx.from_pandas_edgelist(edges[edges['relc'] == 2], 'V1', 'V2', edge_attr='w')
G3 = nx.from_pandas_edgelist(edges[edges['relc'] == 3], 'V1', 'V2', edge_attr='w')

```