### 2020-01-21 paper/glossary.rmd
add glossary rmd

---
output: 
  word_document:
    reference_docx: template.docx
  # github_document
author: Alec Robitaille, Quinn M.R. Webber, Julie Turner and Eric Vander Wal
date: "`r Sys.Date()`"
bibliography: ../references/refs.bib
csl: vancouver.csl
---

## Glossary

# References