### 2020-01-21 paper/glossary.rmd
update to glossary

---
output: 
  word_document:
    reference_docx: template.docx
  # github_document
author: Alec Robitaille, Quinn M.R. Webber, Julie Turner and Eric Vander Wal
date: "`r Sys.Date()`"
bibliography: ../references/refs.bib
csl: vancouver.csl
---

## Glossary

**Interlayer edge**:

**Intralayer edge**:

**Monolayer**:

**Social environment**: Social life of an individual (or group of individuals) described by social interactions (Saltz et al. 2016).

**Spatial network**: Networks of locations, which be linked with social networks of animals that move between the locations (Silk et al. 2018.

# References