### 2020-01-22 README.Rmd
fst Rmd README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
<!-- badges: end -->

# The problem and promise of scale in multilayer networks
Alec Robitaille, Quinn M.R. Webber, Julie Turner and Eric Vander Wal

## Info
[Special Column info](https://academic.oup.com/CZ/pages/animal_social_system)

### Deadlines
* ~~Abstract submission: 30 August 2019~~
* Manuscript submission: 31 January 2020
* Printing: June 2020

### Guidelines
[General Instructions page](https://academic.oup.com/cz/pages/General_Instructions)