### 2020-01-23 paper/glossary.rmd
Update glossary.rmd

---
output: 
  word_document:
    reference_docx: template.docx
  # github_document
author: Alec Robitaille, Quinn M.R. Webber, Julie Turner and Eric Vander Wal
date: "`r Sys.Date()`"
bibliography: ../references/refs.bib
csl: vancouver.csl
---

## Glossary

**Connective redundancy**: 1 - (neighborhood / degree) When connective redundancy is 0, all edges on all layers are necessary to preserve the social ties.[QW: FROM ALEC'S WRITING IN THE MS]

**Degree centrality**:

**Degree deviation**:

**Interlayer edge**:

**Intralayer edge**:

**Monolayer**:

**Multidegree**:

**Neighbours**: degree within a network layer.[QW: FROM ALEC'S WRITING IN THE MS]

**Relevance**: neighborhood of specific layer  / neighbourhood of full ml net. This is the proportion of neighbors present on the layer of focus. Comparing across, it tells you which layers are most important for each individual. [QW: FROM ALEC'S WRITING IN THE MS]

**Social environment**: Social life of an individual (or group of individuals) described by social interactions (Saltz et al. 2016).

**Social scale**: The type of social relationship observed that can be explicitly defined and measured in social network analyses (Farine 2015). For example, social interactions where animals come in direct contact (e.g. grooming) constitute the finest social scale, while co-occurrence or social association constitute an intermediate social scale, and home range overlap constitutes a coarse social scale.

**Spatial network**: Networks of locations, which be linked with social networks of animals that move between the locations (Silk et al. 2018.

# References