### 2020-05-25 README.Rmd
update README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# The problem and promise of scale in multilayer networks
Alec Robitaille, Quinn M.R. Webber, Julie Turner and Eric Vander Wal

## TODO
```{bash}
grep -rni 'R/' -e 'TODO' || true;
grep -rni 'scripts/' -e 'TODO' || true;
grep -rni 'paper/manuscript.Rmd' -e 'TODO' || true;
```