### 2020-06-18 paper/supplement.Rmd
fix encoding in rmd

---
title: Supplementary 1 - Considering scale in multilayer animal social networks
author: Alec L. Robitaille^1\*^, Quinn M.R. Webber^2^, Julie W. Turner^1^, Eric Vander Wal^1,2^
date: |
  | ^1^Department of Biology, Memorial University of Newfoundland, Canada
  | ^2^Cognitive and Behavioural Ecology Interdisciplinary Program, Memorial University of Newfoundland 
  | \*Corresponding author: Alec L. Robitaille (robit.alec@gmail.com)
output:
   bookdown::github_document2: default
   # bookdown::word_document2:
   #   reference_docx: template.docx
knit: (function(inputFile, encoding) {
  rmarkdown::render(input = inputFile, encoding = encoding, output_format = "all") })
bibliography: ../references/knitcite.bib
csl: csl.csl
---

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    # warning = FALSE,
    # message = FALSE,
    eval = TRUE,
    out.width = '\\linewidth'
)

# source('../scripts/00-variables.R')

library(data.table)
DT <- readRDS('../data/derived-data/01-sub-fogo-caribou.Rds')
nobs <- readRDS('../data/derived-data/05-number-of-observations.Rds')
timedefs <- fread('../data/supp-data/timecut-defs.csv')
```

(ref:cutdefs) Time cut definitions

```{r cutdefs}
knitr::kable(timedefs, caption = '(ref:cutdefs)')
```

(ref:lccount) lc count by res

```{r lccount, fig.cap = '(ref:lccount)'}
knitr::include_graphics('../graphics/supp-count-lc.png')
```

(ref:nid) n individuals by time

```{r nid, fig.cap = '(ref:nid)'}
knitr::include_graphics('../graphics/supp-temp-nid.png')
```