### 2020-08-28 README.Rmd
add doi badge

---
output: github_document
---

[![DOI](https://zenodo.org/badge/291094231.svg)](https://zenodo.org/badge/latestdoi/291094231)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# The problem and promise of scale in multilayer networks

* Authors:
    + [Alec L. Robitaille](http://robitalec.ca)
    + [Quinn M.R. Webber](https://qwebber.weebly.com/)
    + [Julie W. Turner](https://www.julwturner.com/)
    + [Eric Vander Wal](http://weel.gitlab.io)