### 2020-09-23 README.Rmd
update README

---
output: github_document
---

[![DOI](https://zenodo.org/badge/291094231.svg)](https://zenodo.org/badge/latestdoi/291094231)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# The problem and promise of scale in multilayer animal social networks

* Authors:
    + [Alec L. Robitaille](http://robitalec.ca)
    + [Quinn M.R. Webber](https://qwebber.weebly.com/)
    + [Julie W. Turner](https://www.julwturner.com/)
    + [Eric Vander Wal](http://weel.gitlab.io)

This repository contains the code accompanying the paper [“The problem and
promise of scale in multilayer animal social networks”](https://doi.org/10.1093/cz/zoaa052). Scripts are under
`scripts/` and reused functions are in `R/`. This project uses standard R
package structure and can therefore be installed with `devtools`. This also
helps declare external package dependencies required for the analysis. Please
note that while functions are included here, they are not tested for use in
other projects and may not be suitable (at least not in their current version).
The project uses `drake` for workflow management and `renv` for explicitly
declaring package dependencies and their versions.

## Abstract
Scale remains a foundational concept in ecology. Spatial scale, for instance,
has become a central consideration in the way we understand landscape ecology
and animal space use. Meanwhile, scale-dependent social processes can range from
fine-scale interactions to co-occurrence and overlapping home ranges.
Furthermore, sociality can vary within and across seasons. Multilayer networks
promise the explicit integration of the social, spatial and, temporal contexts.
Given the complex interplay of sociality and animal space use in heterogeneous
landscapes, there remains an important gap in our understanding of the influence
of scale on animal social networks. Using an empirical case study, we discuss
ways of considering social, spatial and, temporal scale in the context of
multilayer caribou social networks. Effective integration of social and spatial
processes, including biologically meaningful scales, within the context of
animal social networks is an emerging area of research. We incorporate
perspectives that link the social environment to spatial processes across scales
in a multilayer context.